---
title: "data_manipulation_class5"
author: "Yingyu Cui"
date: "2024-09-23"
output: github_document
---
```{r set up}
library(tidyverse)
```
# Import the litter data
```{r import}
litter_df = read_csv("./data_import_examples/FAS_litters.csv")
litter_df = janitor::clean_names(litter_df)
```
# 'Select' some columns but not others

different usages of 'select' function:
- select columns by name
```{r}
select(litter_df, group, litter_number)
select(litter_df, -group)
select(litter_df, group:gd18_weight)
```
- Rename columns:
one  change and select, the other one change but keep others.
```{r}
select(litter_df, GROUP = group, LITTER_NUMBER = litter_number)

rename(litter_df, GROUP = group, LITTER_NUMBER = litter_number)
```
- select columns by condition / select helpers
the last two examples represent the same function but the different expressions.
```{r}
select(litter_df, starts_with("gd"))
select(litter_df, litter_number, everything())
relocate(litter_df, litter_number)
```

# 'Filter' function
select is on columns, filter is on rows.
when we want to see whether a condition is valid, we use "==", instead of "=", which is used to assign values.
the last one repreesents choosing the subjects that meet either of the two conditions but not both.
```{r}
filter(litter_df, gd_of_birth == 20)
filter(litter_df, gd_of_birth != 20)
filter(litter_df, !(gd_of_birth == 20))
filter(litter_df, gd_of_birth <= 20)
filter(litter_df, gd_of_birth == 20, gd0_weight > 22)
filter(litter_df, group == "Mod8")
filter(litter_df, group %in% c("Mod8", "Con7"))
```
## 'Mutate'
mutate is used to create new columns or change the context in different columns.
Question: why the diff(diff_weight = 18_weight-0_weight) cannot calculate but Prof.Jeff could?
```{r}
mutate(litter_df,  
       group = str_to_lower(group))
```
## 'Arrange'
which is put the data in order.(Used not that often)
first A and then B
```{r}
arrange(litter_df, pups_born_alive)
arrange(litter_df, pups_born_alive, gd0_weight)
```

## pipe %>% 
pipe is used to connect the functions together.
one more fuction:
```{r}
litter_df_without_missing = drop_na(litter_df, gd0_weight)
```
A %>% B %>% select(only code the things after ','), mutate works the same way.
lm(wt_gain ~ pups_born_alive, data = _)
sometimes, _ is used to represent the data frame.
there are some cases where what you’re piping isn’t going into the first argument. Here, using the placeholder _ is necessary to indicate where the object being piped should go


If you want to export a vector instead of a dataframe, you can use pull function.
You will often filter using comparison operators (>, >=, <, <=, ==, and !=). You may also use %in% to detect if values appear in a set, and is.na() to find missing values. The results of comparisons are logical – the statement is TRUE or FALSE depending on the values you compare – and can be combined with other comparisons using the logical operators & and |, or negated using !.
group == "Con7" & gd_of_birth == 20

A very common filtering step requires you to omit missing observations. You can do this with filter, but I recommend using drop_na from the tidyr package:

drop_na(litters_df) will remove any row with a missing value
drop_na(litters_df, wt_increase) will remove rows for which wt_increase is missing.

Filtering can be helpful for limiting a dataset to only those observations needed for an analysis. However, I recommend against the creation of many data subsets (e.g. one for each group). This can clutter up your workspace, and we’ll see good tools for the analysis of subsets before long.



|> and the %>% difference


