---
title: "data_wrangling_class4"
author: "Yingyu Cui"
date: "2024-09-22"
output: github_document
---

# We normally put all the libraries we need in the first chunk.
```{r set up}
library(tidyverse)
library(readxl)
library(haven)
```

# Read in some data 
find my current working directory
```{r find my current working directory}
getwd()
```
read in data
```{r read in data}
litter_df = read.csv("./data_import_examples/FAS_litters.csv")
litter_df = janitor::clean_names(litter_df)
```

#take a look at the new data 
```{r take a look at the new data}
head(litter_df)
```
here is another way to look at the data, sometimes there are always something wrong with the last few rows.
```{r another look}
tail(litter_df)
```
here are another fonction to look at the data
```{r another another look} 
skimr::skim(litter_df)
```

the view() function is also a good way to look at the data, but we normally do not open that in the markdown file. This is because it will open a new window in the Rstudio, and will cause some trouble when we knit the file. 

# options in the function read_csv. The funtions could be found in "?read_csv", here they are. ATTENTION: never use read.csv, it is not a good function to read in data. 
read_csv(
  file,
  col_names = TRUE,
  col_types = NULL,
  col_select = NULL,
  id = NULL,
  locale = default_locale(),
  na = c("", "NA"),
  quoted_na = TRUE,
  quote = "\"",
  comment = "",
  trim_ws = TRUE,
  skip = 0,
  n_max = Inf,
  guess_max = min(1000, n_max),
  name_repair = "unique",
  num_threads = readr_threads(),
  progress = show_progress(),
  show_col_types = should_show_types(),
  skip_empty_rows = TRUE,
  lazy = should_read_lazy()
)
col_types = () name the types of data by hand instead of letting R guess it.
na = () recognize some missing data and transform it into "NA"

```{r options in the function read_csv}
litter_df = read_csv("./data_import_examples/FAS_litters.csv", skip = 10, col_names = FALSE)
litter_df = read_csv("./data_import_examples/FAS_litters.csv", na = c("NA", "N/A", "999", "."))
```

# Read other formats, pay attantion to the function of "sheet" and "range"
```{r read other formats}
mlb_df = read_excel("./data_import_examples/mlb11.xlsx", range = "A1:F7")
```
# Read a SAS format file
```{r read a SAS format file}
pulse_df = read_sas("./data_import_examples/pulse.sas7bdat")
pulse_df
``` 
# Use read_csv, rather than use read.csv !!! imp.!!

# Export data 
```{r export data}
write_csv(mlb_df, "./data_import_examples/mlb11_subtable.csv")
```


